Supplementary Information

### 2021-02-17 doc/manuscript/ForestGEO_treerings.Rmd
some changes from Jim Lutz

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF; Figs. `r all_recruitment_plots`), persistent differences in growth rates among individuals could produce a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a species undergoes a dramatic decline in recruitment within a stand (e.g., *Quercus* spp. at SCBI; *Abies alba* at ZOF), such that fast-growing individuals disappear from progressively larger $DBH$ classes as time proceeds. 
In secondary stands, particularly those where many of the sampled species recruited in pulses that were followed by low recruitment (e.g., SCBI, HF; Appendix `r site_details_appendix`, Figs. `r all_recruitment_plots`), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @goulden_patterns_2011; @pregitzer_carbon_2004].
Even within older forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known major disturbance within the past 1.5 centuries (ZOF, CB),  mixed growth trends probably reflect some combination of successional changes and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@vrska_european_2009].

>>>>

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF; Figs. `r all_recruitment_plots`), persistent differences in growth rates among individuals could produce a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a species undergoes a dramatic decline in recruitment within a stand (e.g., *Quercus* spp. at SCBI; *Abies alba* at ZOF), such that fast-growing individuals disappear from progressively larger $DBH$ classes as time proceeds. 
In secondary stands, particularly those where many of the sampled species recruited in pulses that were followed by low recruitment (e.g., SCBI, HF; Appendix `r site_details_appendix`, Figs. `r all_recruitment_plots`), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @goulden_patterns_2011; @pregitzer_carbon_2004].
Even within older forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known major disturbance within the past 1.5 centuries (ZOF, CB),  mixed growth trends probably reflect some combination of successional changes and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@furniss_reconciling_2017; @vrska_european_2009].
